---
title: "5 Census Regressions"
author: "Dan Olner"
date: "5 July 2017"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, comment = NA)
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, warning = F, message = F}
#Preload libraries for myself
library(readr)
#for resizing images
#http://stackoverflow.com/questions/15625990/how-to-set-size-for-local-image-using-knitr-for-markdown
library(png)
library(grid)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(scales)
library(tidyr)

options(scipen=2)

#source("Function_CoBRegressionFunctions.R")

sheet5 <- readRDS('R_data/sheet5.rds')

#Only four cobs
sheet5sub <- readRDS('R_data/sheet5sub.rds')

#summary(lm(data = inputdata, formula = lmfunction)),

#lapply(split(inputdata,inputdata[,splitByFactor]), function(x) summary(lm(data=x, formula = lmfunction)))

#Separate city sheets
citiesList <- c('Glasgow','Edinburgh','Aberdeen','Dundee')

citySheets5 <- readRDS('R_data/citySheets5.rds')
citySheets5sub <- readRDS('R_data/citySheets5sub.rds')
```
\clearpage

#2011 ~ 1971, largest set of 5 census CoBs

* Rich CoBs: Irish Republic, Old Commonwealth, Europe
* Poor: Africa New Commonweath, India, Pakistan, South-East Asia New Commonwealth, Caribbean , Other New Commonwealth, Rest of World

##All zones, rich vs poor

Formula: *xij2011 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71*

```{r echo = F, comment=NA}
lapply(split(sheet5,sheet5[,'rich']), function(x) summary(lm(data=x, formula = xij2011 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre1971)))
```

##Urban zones, rich vs poor

```{r echo = F, comment=NA}
lapply(split(sheet5[sheet5$urbanFiftyPercentPlus==1,],sheet5[sheet5$urbanFiftyPercentPlus==1,'rich']), function(x) summary(lm(data=x, formula = xij2011 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71)))
```

*****

#Four cities (urban zones): rich vs poor, 2011 ~ 1971

```{r echo = F, comment=NA}

for(i in 1:4){
  
print(citiesList[[i]])
  
print(  
lapply(split(citySheets5[[i]],citySheets5[[i]][,'rich']), function(x) summary(lm(data=x, formula = xij2011 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71)))
)

}
```

#Smaller list of CoBs for rich/poor: europe, old commonweath for rich and India, Pakistan for poor

##All zones, rich vs poor (four CoBs)

Formula: *xij2011 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71*

```{r echo = F, comment=NA}
lapply(split(sheet5sub,sheet5sub[,'rich']), function(x) summary(lm(data=x, formula = xij2011 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71)))
```

##Urban zones, rich vs poor (four CoBs)

```{r echo = F, comment=NA}
lapply(split(sheet5sub[sheet5sub$urbanFiftyPercentPlus==1,],sheet5sub[sheet5sub$urbanFiftyPercentPlus==1,'rich']), function(x) summary(lm(data=x, formula = xij2011 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71)))
```

##All zones, rich vs poor (1991 ~ 1971) (four CoBs)

Formula: *xij1991 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71*

```{r echo = F, comment=NA}
lapply(split(sheet5sub,sheet5sub[,'rich']), function(x) summary(lm(data=x, formula = xij1991 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71)))
```

##Urban zones, rich vs poor (1991 ~ 1971) (four CoBs)

```{r echo = F, comment=NA}
lapply(split(sheet5sub[sheet5sub$urbanFiftyPercentPlus==1,],sheet5sub[sheet5sub$urbanFiftyPercentPlus==1,'rich']), function(x) summary(lm(data=x, formula = xij1991 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71)))
```


##Four cities (urban zones), rich vs poor (four CoBs)

```{r echo = F, comment=NA}

for(i in 1:4){
  
print(citiesList[[i]])
  
print(  
lapply(split(citySheets5sub[[i]],citySheets5sub[[i]][,'rich']), function(x) summary(lm(data=x, formula = xij2011 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71)))
)

}
```

#Different decade comparisons 1: 91 vs 71

##All zones, rich vs poor: 1991 ~1971

Formula: *xij1991 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71*

```{r echo = F, comment=NA}
lapply(split(sheet5,sheet5[,'rich']), function(x) summary(lm(data=x, formula = xij1991 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71)))
```

##Urban zones, rich vs poor: 1991 ~ 1971

```{r echo = F, comment=NA}
lapply(split(sheet5[sheet5$urbanFiftyPercentPlus==1,],sheet5[sheet5$urbanFiftyPercentPlus==1,'rich']), function(x) summary(lm(data=x, formula = xij1991 ~ xij1971 + w1971q + migshareMinusOwn1971 + ea1971 + popPerAcre71)))
```

#Different decade comparisons 2: 11 vs 91

##All zones, rich vs poor (11 ~ 91) (four CoBs)

Formula: *xij2011 ~ xij1991 + w1991q + migshareMinusOwn1991 + ea1991 + popPerAcre91*

```{r echo = F, comment=NA}
lapply(split(sheet5sub,sheet5sub[,'rich']), function(x) summary(lm(data=x, formula = xij2011 ~ xij1991 + w1991q + migshareMinusOwn1991 + ea1991 + popPerAcre91)))
```

##Urban zones, rich vs poor (11 ~ 91) (four CoBs)

```{r echo = F, comment=NA}
lapply(split(sheet5sub[sheet5sub$urbanFiftyPercentPlus==1,],sheet5sub[sheet5sub$urbanFiftyPercentPlus==1,'rich']), function(x) summary(lm(data=x, formula = xij2011 ~ xij1991 + w1991q + migshareMinusOwn1991 + ea1991 + popPerAcre91)))
```

##Cities, rich vs poor (11 ~ 91) (four CoBs)

```{r echo = F, comment=NA}

for(i in 1:4){
  
print(citiesList[[i]])
  
print(  
lapply(split(citySheets5sub[[i]],citySheets5sub[[i]][,'rich']), function(x) summary(lm(data=x, formula = xij2011 ~ xij1991 + w1991q + migshareMinusOwn1991 + ea1991 + popPerAcre91)))
)

}
```






